---
title: "Pre-processing cropland datasets"
author: "Lyndon Estes"
date: "29 September 2014"
output:
  html_document:
    theme: spacelab
    toc: yes
---

# Pre-processing reference cropland data layer
<!-- This analysis was undertaken to determine the bias in satellite-derived landcover products at different
spatial scales, using GeoTerraImage's cropland maps for South Africa to assess this error.-->
The following describes the initial pre-processing of landcover datasets that was
undertaken for an analysis of error in landcover data. 

## Gridding the SA landcover data
Much of the original analysis was undertaken on the Mapping Africa server, using postgis/postgres as well as pprepair to do much of the cleaning and conversion of the GTI (reference) dataset into a 1 km resolution cropland percentage cover dataset. The following steps were taken: 

### Cleaning up fields in GTI data
This was done for both the 2007 and 2011 datasets, using a script that pulled out the fields, found the bad ones, read them into an R data object, ran pprepair over it, and then stuck the cleaned geometries back into a postgis table holding the full dataset. The intersection was then performed. 

```{r, eval = FALSE}
source("R_ext/grid-fields-intersect-postgis.R")  # run on mapper.princeton.edu
```

### Calculating field areas 

The next step was running the following script to calculate the area for different cropcover classes in each 1 km^2^ pixel.  This created two core gridded datasets, cover2007.tif and cover2011.tif. 

```{r, eval = FALSE}
source("R_ext/grid-fields-intersect-analyze.R")  # run on mapper.princeton.edu
```



