---
title: 12. Miscellaneous additional analyses

author: "Lyndon Estes"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 4
    number_sections: yes
    pandoc_args: [
      "--number-sections",
      "--number-offset=12"
    ]
vignette: >
  %\VignetteIndexEntry{c12-misc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

\setcounter{section}{12}
<!--# Miscellaneous additional analyses -->

A few additional calculations and figures undertaken for the paper, including supplemental figure (requested by final review) of example illustration of different land-cover datasets analyzed
```{r, message=FALSE}
library(croplandbias)
library(rgeos)
library(dismo)
library(sf)
library(viridis)

rm(list = ls())
p_root <- fp(proj_root("croplandbias"), "croplandbias")
p_edat <- fp(p_root, "external/ext_data")
p_fig <- fp(p_root, "inst/paper/figures/")
```

## South Africa's share of sub-Saharan Africa's area
```{r, echo=FALSE}
# set this to true when want to run all chunks when not knitting
e <- FALSE
```
```{r, message = FALSE, warning = FALSE}
eco <- readOGR(dsn = fp(p_edat, "africa_ecofloristic_zones.sqlite"), 
               layer = "africa_ecofloristic_zones", verbose = FALSE)
crs(eco) <- crs(projection(raster(nrow = 1, ncol = 1)))
data(africa)
ecoalb <- spTransform(eco, crs(africa))

# cut down countries
nms <- africa$cntry_name[africa$region == "Northern Africa"]
nms <- nms[nms != "Sudan"]  # keep Sudan
ssa <- africa[!africa$cntry_name %in% nms, ]
ssa <- ssa[-grep("Island|Sao|Verde|Mauritius|Helena|Comoros|Portugal|Seychell",
                 as.character(ssa@data$cntry_name)), ]
ssa <- gBuffer(ssa, width = 0)
# plot(ssa)

desert <- ecoalb[grep("desert", ecoalb$gez_term), ]
# plot(ecoalb[names(which.max(gArea(desert, byid = TRUE) / 10000 / 100)), ])
sahara <- as.numeric(names(which.max(gArea(desert, byid = TRUE))))

nosahara <- ecoalb[-sahara, ]
# plot(nosahara, col = "red")

ovs <- which(!is.na(over(nosahara, ssa)))
ssaeco <- nosahara[ovs, ]
ssaeco <- ssaeco[-1, ]  # drop a couple Saharan mountains
ssaeco <- ssaeco[-1, ]
# plot(ssaeco[ssaeco@data$gez_term == "Tropical mountain system", ][1, ])
# plot(ssa)
# plot(ssaeco, col = "red", add = TRUE)

# calculate area of remainder
afarea <- gArea(ssaeco) / 10000
# plot(af[af$cntry_name == "South Africa", ])
saarea <- gArea(africa[africa$cntry_name == "South Africa", ]) / 10000
saarea / afarea

```

## Calculation of error propagation
```{r}
load(spathfunc("cropland-werr-2011.rda"))
load(spathfunc("yield-prod-accbias.rda", "yl"))
load(spathfunc("et-err-bias.rda", "et"))
load(spathfunc("carbon-werr.rda", "cb"))

cland <- bacc2011
yprod <- as.data.table(out_tab)[Region == "Density"]
carb <- rbind(cb_statsw_mu, cb_statsw_mua)

# Reshape error stats
# yield/production 
yp1 <- melt(yprod, id.vars = c("Map", "Metric", "Variable"), 
            measure.vars = names(yprod)[5:10])
yp2 <- dcast(yp1, formula = Metric + Variable + variable ~ Map)
setcolorder(yp2, neworder = c(1:3, ncol(yp2), 5:6, 4))
setnames(yp2, names(yp2), c("stnm", "var", "ol", names(cland)[-c(1:2)]))

snms <- names(cland)[-c(1:2)]
lev <- cland[stnm == "mu", ol]

# Production and yield propagation
yp_prop <- rbindlist(lapply(c("Yield", "Production"), function(x) {  
  o <- data.frame(t(sapply(snms, function(y) {  # y <- "sa30"
    emag <- yp2[var == x & stnm == "MAE", get(y)] / 
      cland[stnm == "mua", get(y)] 
    c(emag, mean(emag))
  })))
  o <- cbind(snms, round(o, 2))
  colnames(o) <- c("map", lev, "mu")
  data.table(cbind.data.frame("var" = x, o))
}))

# Carbon differences
c_prop <- rbindlist(lapply(snms, function(x) {  # x <- "sa30"
  dat <- carb[stnm == "MAE" & map == x, names(carb)[-c(1:3)], with = FALSE]
  o <- dat[, lapply(.SD, function(y) y / cland[stnm == "mua", get(x)]), 
           .SDcols = names(dat)]
  o <- round(rbind(o, data.table(t(colMeans(o)))), 2)
  o <- cbind("ol" = c(lev, "mu"), o)
  cbind("map" = x, o)
  # dat[, second] / cland[stnm == "mua", get(x)]
}))

# et propagation
et_prop <- round(et_err[Variable == "Annual Mean", MAE] / 
  cland[ol == "f25" & stnm == "mua", snms, with = FALSE], 2)

yp_prop
c_prop
et_prop
```

## Land cover data illustrated
### Read in base land cover in Albers format

For this part, steps taken to get to Albers-transformed versions of base land cover data subset to South Africa's extent will have to be repeated. Due to size, these were not saved in the repo. See [chapter 1](c1-landcover-fetch.html) and [chapter 3](c3-landcover-preprocess-2.html) to recreate these initial steps. 
```{r, eval = e}
sa_r <- raster(spathfunc("sagrid.tif"))
glc <- raster(fp(p_edat, "landcover/glc_share/glcSA_alb_rect.tif"))
glob <- raster(fp(p_edat, "landcover/globcover_2009/globSA_alb.tif"))
modis <- raster(fp(p_edat, "landcover/MCD12Q1/modis_type_1_alb.tif"))
salc <- raster(fp(p_edat, "landcover/sa_lc_2009/sa_lc_2009.tif"))  

# Also read in raw shapefile of South Africa field boundaries, for illustration
gti <- st_read(dsn = fp(p_edat, "landcover/gti/GTI_SA_2011.shp"), 
               layer = "GTI_SA_2011")

# small image of South Africa, in GCS
ext <- extent(c(26.95, 27.35, -28.72, -28.43))  # arbitrary small cropland area
base <- gmap(ext, type = "satellite", lonlat = TRUE, rgb = TRUE)
base <- trim(base)
baseext <- as(extent(base), "SpatialPolygons")
gcs <- data.table(make_EPSG())[like(code, "4326"), prj4]
crs(baseext) <- CRS(gcs)
baseext_alb1 <- spTransform(baseext, crs(sashp))  # albers used for paper
baseext_alb2 <- spTransform(baseext, crs(salc))  # albers used for native SA-LC

```

### Crop them down to save
```{r, eval = e}
gtic <- st_intersection(gti, st_as_sf(baseext))  # crop GTI layers
salcc <- crop(salc, baseext_alb2)  # crop with native albers extent
globc <- crop(glob, baseext_alb1)  # crop with albers used in paper
modisc <- crop(modis, baseext_alb1)  # ""
glcc <- crop(glc, baseext_alb1)  # ""

# Set minor (<10 pixels) cover types to NA
# hist(salcc, maxpixels = ncell(salcc), plot = FALSE, breaks = seq(0.5, 5.5, 1))
# only 6 pixels in degraded in degraded class, so plays havoc with levelplot and 
# ggplot--set to NA
salcc[salcc == 3] <- NA

# table(values(modisc))  # class 9 and 13 only have 2 and 6 each. Set both to NA
modisc[modisc %in% c(9, 13)] <- NA

# table(values(globc))  # class 14 (rainfed cropland), 60, and 130 have 1, 8, 2
globc[globc %in% c(14, 60, 130)] <- NA

# add categories to categorical data
# cols <- rev(terrain.colors(8))
# cols <- c("khaki")

# display.brewer.all()
crcols <- brewer.pal(11, "PuOr")[3:5]
blth20 <- brewer.pal(11, "Paired")[6]
natveg <- brewer.pal(9, "YlGn")[c(2, 3, 4)]
# plot(1:5, pch = 20, col = natveg, cex = 3)

# SA-LC
salcc <- ratify(salcc)
rat <- levels(salcc)[[1]]
rat[["lc"]] <- c("Natural", "Cultivation", "Built", "Water")
rat[["cols"]] <- c(natveg[3], crcols[1], blth20, "blue")
levels(salcc) <- rat

# GlobCover
globc <- ratify(globc)
rat <- levels(globc)[[1]]
rat[["lc"]] <- c("Mosaic veg/crop", "Mosaic for-shrub/grass", 
                 "Closed/open herb", "Sparse veg", "Water")
rat[["cols"]] <- c(crcols[2], rev(natveg), "blue")
levels(globc) <- rat

# MODIS
modisc <- ratify(modisc)
rat <- levels(modisc)[[1]]
rat[["lc"]] <- c("Open shrub", "Grasslands", 
                 "Croplands", "Cropland/nat. veg")
rat[["cols"]] <- c(natveg[2:1], crcols[1], crcols[2])
levels(modisc) <- rat


```

### And plot
```{r, eval = e}
sasf <- st_as_sf(spTransform(sashp, CRS(gcs)))  # sasp to sf (gcs)

l <- -1
cx <- 0.8
png(fp(p_fig, "lc-demo-fig.png"), width = 6, height = 2.5, units = "in", 
    res = 300)
par(mfrow = c(2, 4), mar = c(0, 0, 0, 1.5), oma = c(0, 0, 0, 0))

# sa shp
plot(st_geometry(sasf), col = "grey")
plot(st_as_sf(baseext), col = "red", add = TRUE)

# Base image
plot(baseext, lty = 0)
plotRGB(base, add = TRUE)

# Base image with GTI polys
plot(baseext, lty = 0)
plotRGB(base, add = TRUE)
mtext("Reference", side = 3, line = l, cex = cx)
plot(st_geometry(gtic), border = "transparent", col = crcols[1], add = TRUE)

# GLC-Share
aargs <- list(mgp = c(1, 0.25, 0), at = seq(0, 100, 10), 
              tcl = -0.2, cex.axis = 0.6)
plot(baseext_alb1, lty = 0)
plot(glcc, add = TRUE, axes = FALSE, legend = FALSE, box = FALSE, 
     interpolate = FALSE, col = inferno(10), breaks = seq(0, 100, 10))
mtext("GLC-Share", side = 3, line = l, cex = cx)
plot(glcc, legend.only = TRUE, title = "%", legend.width = 1, col = inferno(10),
     legend.shrink = 0.5, axis.args = aargs, 
     legend.args = list("%", cex = 0.6))

# SA-LC, GlobCover, MODIS
plot(baseext_alb2, lty = 0)
plot(salcc, add = TRUE, axes = FALSE, legend = FALSE, box = FALSE,
     col = levels(salcc)[[1]]$col)  # SA-LC
mtext("SA-LC", side = 3, line = l, cex = cx)
plot(baseext_alb1, lty = 0)
plot(globc, add = TRUE, axes = FALSE, legend = FALSE, box = FALSE, 
     col = levels(globc)[[1]]$col)  # Glob-Cover
mtext("GlobCover", side = 3, line = l, cex = cx)
plot(baseext_alb1, lty = 0)
plot(modisc, add = TRUE, axes = FALSE, legend = FALSE, box = FALSE, 
     col = levels(modisc)[[1]]$col)  # MODIS
mtext("MODIS", side = 3, line = l, cex = cx)

# Legend
legtxt <- c(" NAT / FSG /  -  ", 
            "  -  / COH / OSH ", 
            "  -  / SVG / GRL ",
            " CUL /  -  / CLD", 
            "  -  / MVC / CNV ", 
            " BLT /  -  /  - ",
            " H20 / H20 /  -  ")
plot(baseext, lty = 0)
par(family = "mono")
legend("left", bty = "n", legend = legtxt, title = "SA-LC/GlobCover/MODIS",
       fill = c(rev(natveg), crcols[1:2], blth20, "blue"), 
       inset = -0.005, y.intersp = 0.7, x.intersp = 0.5, cex = 0.8, 
       title.adj = 3.5)

dev.off()







```



