---
title: "2. Pre-processing reference cropland data"
author: "Lyndon Estes"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 4
    number_sections: yes
    pandoc_args: [
      "--number-sections",
      "--number-offset=2"
    ]
vignette: >
  %\VignetteIndexEntry{c2-landcover-preprocess-1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

\setcounter{section}{2}
<!--# Pre-processing reference cropland data layer-->
The following describes the initial pre-processing of landcover datasets that was undertaken for an analysis of error in landcover data. 

Much of the original analysis was undertaken on the Mapping Africa server, using postgis/postgres as well as pprepair to do much of the cleaning and conversion of the GTI (reference) dataset into a 1 km resolution cropland percentage cover dataset. The following steps were taken: 

## Cleaning up fields in GTI data
This was done for both the 2007 and 2011 datasets, using a script that pulled out the fields, found the bad ones, read them into an R data object, ran pprepair over it, and then stuck the cleaned geometries back into a postgis table holding the full dataset. The intersection was then performed. 

```{r, eval = FALSE}
source("grid-fields-intersect-postgis.R")  # run on mapper.princeton.edu
```

## Calculating field areas 

The next step was running the following script to calculate the area for different cropcover classes in each 1 km^2^ pixel.  This created two core gridded datasets, cover2007.tif and cover2011.tif. 

```{r, eval = FALSE}
source("grid-fields-intersect-analyze.R")  # run on mapper.princeton.edu
```

Note: both scripts are available as system files under the sub-folder, and can be found using the following commands. 
```{r, eval = FALSE}
library(croplandbias)
system.file("extR", "grid-fields-intersect-postgis.R", 
            package = "croplandbias")
system.file("extR", "grid-fields-intersect-analyze.R", 
            package = "croplandbias")
```




